---
title: "Section B"
output: html_notebook
---


# Clearing the environment
```{r}
rm(list = ls())

```


```{r}
Dataset_Raw = load("section_b.Rda")            #  -> Reading Rda


```


1.
```{r}
#names(sales)
TotalSoda = sum(sales$soda)

print(TotalSoda)
```


2.
```{r}
max(sales$wraps)

```


3.
```{r}
mean(sales$total_sales[sales$staff == "Harry"])
```

4.
```{r}
mean(sales$total_sales[sales$staff == "Sara"])

```


5.
```{r}

Harry_Sales = sales$total_sales[sales$staff == "Harry"]
Sara_Sales  = sales$total_sales[sales$staff == "Sara"]
t.test(Harry_Sales,Sara_Sales, mu = 0,  alt = "two.sided", conf = 0.95 )
```

6.
At a significance of 0.95 the alpha is 0.05, since we are doing a two sided test and the p-value is 0.06028, which is higher than the alpha, we don`t reject the null hypothesis that the average difference in sales made by Harry and Sara is 0

7.
```{r}
MaxHumilty = max(sales$humidity)
MinHumilty = min(sales$humidity)

print(MaxHumidity)

print(MinHumidity)
```

8.
```{r}
attach(sales)
model1 = lm(coffee ~ temp)
model2 = lm(coffee ~ humidity)
model3 = lm(coffee ~ wind)
detach(sales)
```

9.
```{r}
screenreg(list(model1,model2,model3))
```

10.
As we can see from the resuylts of the regression, the coefficient for "temp" is significant and negative, that means than on average a 1 point increase in the variable temp means a 0.59 point decrease in the variable coffee, this relationship is significant to a confidece of 99.9%

The coefficient for humidity is significant with a level of 99.9% and is positive, that means that on average a 1 point increase in the variabl;e humidity will be reflected as a 0.4 increase in the coffee variable.

The coefficient for the variable wind is non significant at a 95% confidence, that means that the coefficient is statistically equal to 0, and there is no linear relationship between the variable wind and the variable coffee.

11.
```{r}
# Humidity = c(30,80)
dataframe_ToPredict = data.frame(humidity = c(30,80))

predict(model2,dataframe_ToPredict)
```
We can see from the results that the predicted coffees sold when the humidity is 30 is 10.32037, when the humidity is 80 is 30.29006

If we wanted we could have specified also the levels for the other variable in the dataframe.







sales